---
title: "Analysis plots for FRC paper"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2) # for all plots
library(dplyr) # data wrangling
library(purrr) # efficient loading of files
library(here) # handle file locations
library(knitr) # make an html
library(ggsignif) # call signifance in ggplot
library(patchwork) # put plots together
library(wesanderson) # shiny colors
```

##### *Code by Friedrich Preusser (github.com/Fritze)*
##### All organoids acquired with Z1 lightsheet (August 2019).
##### Data shown here is all from *Draq5* channel.
###### --- 
###### 1. Multiview recontruction based on Draq5 signal (nuclei as interest points).
###### 2. 400x400 bounding box through full z (for confocal: 800x800 px).
###### 3. FRC-QE, **FFT size 200**, **stepsize=1** 
###### 4. For all measurements: Measurement per slice.

### Select a tab to view it's content   {.tabset}

#### Code to import data
```{r, echo=TRUE, warning=FALSE}

#to run properly the script has to be in a folder that contains the folder "data" with all quantifications
list_data_csvs <- list.files(here("data"),".+csv$",include.dirs= FALSE)

read_data <- function(files){
  read.csv(files) %>%
  rename_all(tolower) %>%
  #extract protocol from beginning of file name
  mutate(protocol = gsub(".+\\_(.+)\\_org.+","\\1",basename(files))) %>%
  #extract measurement type from beginning of file name
  mutate(measured = gsub("(^[^_]+).+","\\1",basename(files))) %>%
  #extract OrganoidID (starting with "org" or "orgconf (for confocal data))
  mutate(orgID = gsub(".+\\_(org\\d+|orgconf\\d+).+","\\1",basename(files))) %>%
  #extract filename
  mutate(filename=basename(files)) %>%
  #extract angle
  mutate(angle=gsub(".+angle(\\d+).+","\\1",filename)) %>%
  mutate(angle=ifelse(angle == 0, "180°",angle)) %>%
  mutate(angle=ifelse(angle == 1,"0°",angle)) %>%
  #if angle information is missing in filename make it "both angles"
  mutate(angle = ifelse(grepl("angle",filename),angle,"both_angles")) %>%
  #extract illumination side
  mutate(illu=gsub(".+illu(.+)\\_angle.+","\\1",filename)) %>%
  mutate(illu=ifelse(illu == 0, "B",illu)) %>%
  mutate(illu=ifelse(illu == 1,"A",illu)) %>%
  mutate(illu=ifelse(illu == "fused","both_illuminations",illu)) %>%
  #if illumination information is missing in the filename make it "both_illuminations"
  mutate(illu = ifelse(grepl("illu",filename),illu,"both_illuminations")) %>%
  #extract staining type 
  #if staining information is missing in the filename make it "draq5"
  mutate(staining = ifelse(grepl("mito",filename),"mitotracker","draq5")) %>%
  #all values (either quality or image feature measurements) should be named "value"
  rename_all(~sub('quality', 'value', .x)) %>%
  rename_all(~sub('mean','value',.x)) %>%
  #slice numbering in "slice" column
  mutate(slice = (1:n())-1)

}




data_full <- map_dfr(file.path(here("data"),list_data_csvs),read_data) %>% 
  rename(min_Img = min, max_Img = max) %>%
  select(protocol,orgID,slice,angle,illu,staining,measured, value,max_Img,min_Img) %>%
  mutate(measured = ifelse(measured == "Img-features","mean-intensity",measured)) %>%
  group_by(protocol,orgID,illu,angle) %>%
  mutate(max_slice = max(slice)) %>%
  mutate(slice_norm = slice-max_slice/2) %>%
  group_by(protocol,orgID,angle,illu,staining,measured) %>%
  #contrast is just max minus minimum intensity at a given slice
  mutate(contrast = max_Img - min_Img) %>%
  mutate(contrast_norm = contrast / max(contrast)) %>%
  select(-c(max_Img,min_Img)) %>%
  #normalize all measured values between 0 and 1
  mutate(value_norm = value / max(value)) %>% 
  ungroup() %>%
  arrange(protocol,orgID)



stepsizes_and_corresponding_orgIDs <- data.frame(
                    protocol = c("ClearT2","ClearT2","ClearT2",
                                  "Fructose","Fructose","Fructose",
                                  "ScaleA2","ScaleA2","ScaleA2",
                                 "Fructose","Fructose","Fructose",
                                 "Control","Control","Control"),
                    orgID = c("org1","org2","org3",
                              "org2","org4","org5",
                              "org1","org2","org3",
                              "orgconf1","orgconf2","orgconf3",
                              "orgconf1","orgconf2","orgconf3"),
                    stepsize = c(2.42, 2.42, 2.42,
                                1.7,1.45,1.55,
                                2.42,2.42,2.42,
                                2,2,2,
                                2,2,2))

data_full <- data_full %>%
  left_join(stepsizes_and_corresponding_orgIDs,by=c("protocol","orgID")) %>%
  mutate(slice_norm_micro = slice_norm * stepsize)


```


#### Plot for APPLICATION NOTE Figure 1
```{r, echo=TRUE, warning=FALSE,dpi = 600,fig.height=5,fig.width=10}

#select organoid to plot
selected_organoid <- "ScaleA2_org2"
boundaries <- c(15,70,210,260)

# selected_slices <- c(7,60,134,220)


boundaries_micron <- filter(data_full,slice %in% boundaries) %>%
  #make name
  mutate(name = paste0(protocol,"_",orgID)) %>%
  #only the organoid we want to plot
  filter(name == selected_organoid) %>%
  distinct(slice_norm_micro) %>%
  pull(slice_norm_micro)

#dataframe describing edge locations
rects <- data.frame(xmin = c(boundaries_micron[1],boundaries_micron[2],boundaries_micron[3]),
                    xmax = c(boundaries_micron[2],boundaries_micron[3],boundaries_micron[4]),
                    ymin = -Inf,  
                    ymax = Inf,
                    fill = c( wes_palette("Moonrise3")[3], wes_palette("Moonrise3")[1], wes_palette("Moonrise3")[3]))

#filter data for plots
data_to_plot <- data_full %>%
  #only draq5 staining
  filter(staining == "draq5") %>%
  #only multiview reconstructed images
  filter(angle == "both_angles") %>%
  #make name
  mutate(name = paste0(protocol,"_",orgID)) %>%
  #only the organoid we want to plot
  filter(name == selected_organoid)  %>%
  filter(measured %in% c("mean-intensity", "DCT-Shannon-entropy","FRC"))

all_plot <- ggplot(data_to_plot, aes_string(x="slice_norm_micro",y="value_norm",color="measured"))+
  # geom_rect(data = rects, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = fill), inherit.aes = FALSE, alpha = 0.15) +
  geom_line(aes(color=measured),size=1.5)  +
  geom_vline(xintercept=rects$xmin, size=0.3,alpha = 0.15) +
  geom_vline(xintercept=rects$xmax, size=0.3,alpha = 0.15) +
  theme_bw() +
  labs(y = "value (normalized)")+
  scale_x_continuous(limits = c(min(data_to_plot$slice_norm_micro),max(data_to_plot$slice_norm_micro)),
                     breaks = c(-300,-200,-100,0,100,200,300))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        plot.title = element_text(size=10))





#plot
plot <- all_plot & theme(panel.border = element_rect(size=2),
              axis.ticks = element_line(colour = "black", size = 1.5),
              axis.text = element_text(face="bold",size=12),
              panel.background = element_rect(fill = "transparent"), # bg of the panel
              plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
              panel.grid.major = element_blank(), # get rid of major grid
              panel.grid.minor = element_blank(), # get rid of minor grid
              ) &
        scale_color_brewer(palette = "Set2")


ggsave(plot=plot, file="AN_Figure_1_lineplots.png", type = "cairo-png",  bg = "transparent",dpi = 1800,height=3,width=10)

plot
```


#### Supplemental Figure 4 - Comparing illuminations
```{r, echo=TRUE, warning=FALSE,dpi = 300,fig.height=4,fig.width=8}

#define line plot function for the following figures
line_plot <- function(data_to_plot,X,Y,xlab,ylab,colorby){
  ggplot(data_to_plot, aes_string(x=X,y=Y,group="name"))+
    geom_line(aes_string(color=colorby),size=1) + 
    labs(x=xlab, y=ylab) +
    facet_grid(~protocol,scale="free_x")+
    theme_bw() +
    theme(legend.position = "top")
}

data_to_plot <- data_full %>%
  filter(measured %in% c("FRC-ROI-left","FRC-ROI-right")) %>%
  mutate(name = paste(measured,illu, sep="_")) %>%
  filter(slice_norm_micro < 100 & slice_norm_micro > -100)

FRC_ROIs_plot <- data_to_plot %>%
  line_plot(.,"slice_norm_micro","value","µm (relative to center)","FRC-QE score","illu")+
  facet_wrap(~measured,scales="free")


patch <- FRC_ROIs_plot


#plot
plot <- patch & 
    theme(panel.border = element_rect(size=1.5),
              axis.text = element_text(face="bold",size=12),
              strip.text = element_text(face="bold",size=14)
              ) &
    scale_color_manual(values = wes_palette("Moonrise3"))

ggsave(plot=plot, file="S4_illuminations.png", type = "cairo-png",  bg = "transparent",dpi = 300,height=4,width=8)

plot


```


#### Supplemental Figure S5 - Multi-view quality estimation
```{r, echo=TRUE, warning=FALSE,dpi = 300,fig.height=4,fig.width=8}

unfused_examples <- data_full %>%
  mutate(name = paste0(protocol,"_",orgID)) %>%
  filter(!grepl("both_angles",angle)) %>% 
  distinct(name) %>%
  pull(name)
  
selected_slices_micron <- c(0,200)

data_to_plot <- data_full %>%
  #only draq5 staining
  filter(staining == "draq5") %>%
  mutate(name = paste0(protocol,"_",orgID)) %>%
  filter(name %in% unfused_examples) %>%
  filter(!grepl("ScaleA2",name)) %>% 
  #select specific organoids
  filter(name %in% c("ClearT2_org2", "Fructose_org4")) %>% 
  mutate(name = paste0(protocol,"_",orgID)) %>%
  mutate(microns_from_start = slice * stepsize) %>%
  mutate(angle = ifelse(!grepl("both_angles",angle),angle,"multiview fusion")) %>%
  filter(measured == "FRC")

#fixing the order
data_to_plot$angle_ordered <- factor(data_to_plot$angle, levels=c("0°","multiview fusion", "180°"))

FRC_plot <- line_plot(data_to_plot,"slice_norm_micro","value","µm (relative to center)","FRC-QE score","name") +
  facet_grid(~angle_ordered,scale="free_x") +
  scale_y_continuous(limits=c(0, max(data_to_plot$value)))


#put everything together
patch <-  FRC_plot


#plot
plot <- patch & 
    theme(panel.border = element_rect(size=1.5),
              axis.text = element_text(face="bold",size=12),
              strip.text = element_text(face="bold",size=14)
              ) &
    scale_color_manual(values = wes_palette("Moonrise3")) &
    geom_vline(xintercept=selected_slices_micron,linetype='dashed',size=0.8,color="red") 

ggsave(plot=plot, file="S5_lightsheet.png", type = "cairo-png",  bg = "transparent",dpi = 300,height=4,width=8)

plot

```

#### Supplemental Figure S6 - Confocal microscopy
```{r, echo=TRUE, warning=FALSE,dpi = 300,fig.height=4,fig.width=8}

replicates_and_corresponding_orgIDs <- data.frame(protocol =
                                          c("Fructose","Fructose","Fructose",
                                            "Control","Control","Control"),
                    orgID = c("orgconf1","orgconf2","orgconf3",
                              "orgconf1","orgconf2","orgconf3"),
                    replicates = c("replicate 3", "replicate 2", "replicate 1",
                                  "replicate 3","replicate 2","replicate 1"))

selected_slices_micron <- c(150)


data_to_plot <- data_full %>%
  filter(grepl("conf",orgID)) %>%
  left_join(replicates_and_corresponding_orgIDs,by=c("orgID","protocol")) %>%
  mutate(protocol = ifelse(protocol == "Control", "Uncleared",protocol)) %>%
  mutate(name = paste0(protocol,"_",orgID)) %>%
  mutate(microns_from_start = slice * stepsize)

FRC_plot <- data_to_plot %>%
  filter(measured == "FRC") %>% 
  line_plot(.,"microns_from_start","value","µm inside","FRC-QE score","replicates")



#put everything together
patch <- FRC_plot


#plot
plot <- patch & 
    theme(panel.border = element_rect(size=1.5),
              axis.text = element_text(face="bold",size=12),
              strip.text = element_text(face="bold",size=14)
              ) &
    scale_color_manual(values = wes_palette("Moonrise3"))

ggsave(plot=plot, file="S6_confocal.png", type = "cairo-png",  bg = "transparent",dpi = 300,height=4,width=8)

plot

```



#### Supplemental Figure S7 - Comparing across protocols
```{r, echo=TRUE, warning=FALSE,dpi = 300,fig.height=8,fig.width=16}

selected_slices_micron <- c(-200,0)


replicates_and_corresponding_orgIDs <- data.frame(protocol = c("ClearT2","ClearT2","ClearT2",
                                  "Fructose","Fructose","Fructose",
                                  "ScaleA2","ScaleA2","ScaleA2"),
                    orgID = c("org3","org2","org1",
                              "org5","org4","org2",
                              "org2","org3","org1"),
                    replicates = c("replicate 1", "replicate 2", "replicate 3",
                                  "replicate 1","replicate 2","replicate 3",
                                  "replicate 1","replicate 2","replicate 3"))

  

data_to_plot <- data_full %>%
  left_join(replicates_and_corresponding_orgIDs,by=c("orgID","protocol")) %>%
  #only fused angles
  filter(angle == "both_angles") %>%
  #only both illuminations
  filter(illu == "both_illuminations") %>%
  #only draq5 staining
  filter(staining == "draq5") %>%
  mutate(name = paste0(protocol,"_",orgID)) %>%
  #filter out confocal data
  filter(!grepl("orgconf",orgID))


contrast_plot <- data_to_plot %>%
  filter(measured == "mean-intensity") %>%
  line_plot(.,"slice_norm_micro","contrast","slice","Contrast","replicates")+
    labs(y="Contrast",x="µm (relative to center)")


DCT_Shannon_entropy_plot <- data_to_plot %>%
  filter(measured == "DCT-Shannon-entropy") %>%
  line_plot(.,"slice_norm_micro","value","slice","DCT Shannon entropy","replicates")+
    labs(y="Entropy",x="µm (relative to center)")

FRC_plot <- data_to_plot %>%
  filter(measured == "FRC") %>% 
  line_plot(.,"slice_norm_micro","value","slice","FRC score (raw)","replicates") +
  labs(y="FRC",x="µm (relative to center)")


#put all plots together
patch <- contrast_plot / DCT_Shannon_entropy_plot / FRC_plot 


#plot
plot <- patch & 
    theme(panel.border = element_rect(size=1.5),
              axis.text = element_text(face="bold",size=12),
              strip.text = element_text(face="bold",size=14)
              ) &
    scale_x_continuous(limits= c(min(data_to_plot$slice_norm_micro),max(data_to_plot$slice_norm_micro)),
                     breaks = c(-300,-150,0,150,300)) &   
    scale_color_manual(values = wes_palette("Moonrise3")) &
    geom_vline(xintercept=selected_slices_micron,linetype='dashed',size=0.8,color="red")

ggsave(plot=plot, file="S7_lineplots.png", type = "cairo-png",  bg = "transparent",dpi = 300,height=8,width=16)

plot

```

#### Box Plot for Supplemental Figure 7
```{r, echo=TRUE, warning=FALSE,dpi = 300,fig.height=3.5,fig.width=7}
#make sampling reproducible
set.seed(2020)

#take data_to_plot dataframe from previous chunk
#and select 200 slices per experiment (x 3 replicates x 3 protocols = 1800 slices in total)
selected_slices <- data_to_plot %>%
  #only fused angles
  filter(angle == "both_angles") %>%
  #only both illuminations
  filter(illu == "both_illuminations") %>%
  #only draq5 staining
  filter(staining == "draq5") %>%
  mutate(name_slice = paste0(name,"_",slice)) %>%
  group_by(orgID,protocol) %>%
  sample_n(200) %>%
  pull(name_slice)

#filter out FRC for only those slices
data_to_plot2_FRC <- data_to_plot %>%
  mutate(name_slice = paste0(name,"_",slice)) %>%
  filter(measured == "FRC") %>%
  filter(name_slice %in% selected_slices)

boxplot_FRC <- ggplot(data_to_plot2_FRC, aes(y=value,x=protocol,group=protocol,fill=protocol))+
  geom_violin(fill="lightgrey",alpha=0.8,size=0,inherit_aes=FALSE,outlier.shape=NA)+
  geom_boxplot(width=0.25,size=1,alpha=.3,outlier.alpha = 1) +
  scale_y_continuous(limits=c(0,max(data_to_plot2_FRC$value + 20)))+
  geom_signif(comparisons = list(c("Fructose", "ScaleA2"),c("ClearT2","Fructose")), 
              map_signif_level=TRUE,y_position = c(52, 57),test = "wilcox.test",size=1)+
  geom_signif(comparisons = list(c("ClearT2","ScaleA2")),
              map_signif_level = TRUE,y_position=c(62),test = "wilcox.test",size=1)+
  labs(x="protocol", y="FRC") +
  scale_colour_discrete(name = "Organoid ID") +
  theme_bw() +
  theme(legend.position = "right")+
  scale_fill_brewer(palette = "Set2")+
  coord_flip()

#filter out DCT Shannon entropy values for the same slices
#so that we look at the same images in both plots
data_to_plot2_DCT_Shannon_entropy <- data_to_plot %>%
  mutate(name_slice = paste0(name,"_",slice)) %>%
  filter(measured == "DCT-Shannon-entropy") %>%
  filter(name_slice %in% selected_slices)

boxplot_DCT_Shannon_Entropy <- ggplot(data_to_plot2_DCT_Shannon_entropy, aes(y=value,x=protocol,group=protocol,fill=protocol))+
  geom_violin(fill="lightgrey",alpha=0.8,size=0,inherit_aes=FALSE,outlier.shape=NA)+
  geom_boxplot(width=0.25,size=1,alpha=.3,outlier.alpha = 1) +
  scale_y_continuous(limits=c(0,max(data_to_plot2_DCT_Shannon_entropy$value + 0.007)))+
  geom_signif(comparisons = list(c("Fructose", "ScaleA2"),c("ClearT2","Fructose")),
              map_signif_level=TRUE,y_position = c(0.015, 0.02),test = "wilcox.test",size=1)+
  geom_signif(comparisons = list(c("ClearT2","ScaleA2")),
  map_signif_level = TRUE,y_position=c(0.023),test = "wilcox.test",size=1)+
  labs(x="protocol", y="Entropy") +
  scale_colour_discrete(name = "Organoid ID") +
  theme_bw() +
  theme(legend.position = "right")+
  scale_fill_brewer(palette = "Set2")+
  coord_flip()


plot <- boxplot_FRC & theme(panel.border = element_rect(size=1.5),
              axis.text = element_text(face="bold",size=12),
              legend.direction  = "vertical",
              legend.key.height =unit(2, "cm"),
              legend.key.width = unit(1,"cm"))

ggsave(plot=plot, file="S7_FRC_boxplot.png", type = "cairo-png",  bg = "transparent",dpi = 300,height=3.5,width=7)

plot

plot <- boxplot_DCT_Shannon_Entropy & theme(panel.border = element_rect(size=1.5),
              axis.text = element_text(face="bold",size=12),
              legend.direction  = "vertical",
              legend.key.height =unit(2, "cm"),
              legend.key.width = unit(1,"cm"))

ggsave(plot=plot, file="S7_DCT-Shannon-Entropy_boxplot.png", type = "cairo-png",  bg = "transparent",dpi = 300,height=3.5,width=7)

plot

```


```{r, echo=TRUE, warning=FALSE,dpi = 300,fig.height=8,fig.width=16}

# library(reshape2)
data_to_plot <- data_full %>%
  left_join(replicates_and_corresponding_orgIDs,by=c("orgID","protocol")) %>%
  mutate(name = paste(protocol, orgID, sep="_")) %>%
  filter(protocol == "ScaleA2") %>%
  filter(measured == "BRISQUE" | measured == "FRC") %>%
  filter(angle == "both_angles") %>%
  filter(illu == "both_illuminations") %>%
  filter(slice_norm_micro < 100 & slice_norm_micro > -100)

s <- filter(data_to_plot, measured == "FRC")
max(s$value)

ggplot(data_to_plot, aes(y=value, x=slice_norm_micro, color=replicates))+
  geom_line(size=1.25)+
  facet_wrap(~measured, scales="free_y")+
  theme_classic()+
   scale_color_manual(values = wes_palette("Moonrise3"))+
  expand_limits(y = 0)+
  ggsave(file="FRC-brisque_comparison_ScaleA2.png", type = "cairo-png",  bg = "transparent",dpi = 300,height=4,width=8)


```

